---
title: "Workshop 02. Data Pre-processing"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Package Installation
### Caret
The Caret package has some issues with versions compatibility, so the installation is a bit tricky.
```{r eval = FALSE}
install.packages('caret',dependencies=TRUE)
#DO NOT RUN NOW
update.packages(checkBuilt = TRUE)
```

### Other packages
```{r eval = FALSE}
install.packages('corrplot')
install.packages('e1071')
install.packages('lattice')
```

```{r}
library(caret)
library(corrplot)
library(e1071)
library(lattice)
```

## Load Cell Segmentation Data

```{r}
library(AppliedPredictiveModeling)
```

```{r}
data(segmentationOriginal)
```

```{r}
structure(segmentationOriginal)
```

```{r}
summary(segmentationOriginal)
```
First three columns:
	Cell, Class (PS or WS), Case (training or set)
	
```{r}
segData <- subset(segmentationOriginal, Case == "Train")
```

```{r}
Cell <- (segData$Cell)
Class <- segData$Class   # factor
Case <- segData$Case 
```

Remove first three columns Cell, Class, Case
```{r}
segData <- segData[, -(1:3)]
```

```{r}
statusColNum <- grep("Status", names(segData))
#remove column names status
segData <- segData[, -statusColNum] 
```

## PCA with base R
```{r}
pcaObject <- prcomp(segData,center=TRUE,scale.=TRUE)
```

Explore PCA Object
```{r}
pcaObject 
```

Calculate the cumulative percentage of variance which each component > # accounts for.
```{r}
percentVariance <- pcaObject$sd^2/sum(pcaObject$sd^2)*100
percentVariance[1:3]
```
The transformed values are stored in pcaObject as a sub-object called x: 
```{r}
head(pcaObject$x[, 1:5])
```

```{r}
head(pcaObject$rotation[, 1:3]) 
```

## Pre-processing with caret
- transform
- center 
- scale
- impute values 
- spatial sign transformation 
- feature extraction 

### preProcess from caret
```{r}
trans <- preProcess(segData, method = c("BoxCox", "center", "scale", "pca"))
trans
```

Apply the transformations:
```{r}
transformed <- predict(trans, segData)
head(transformed[, 1:5])
```


### Filtering

Check for near Zero variance
```{r}
nearZeroVar(segData)
```


Correlations between predictor variables
```{r}
nearZeroVar(segData)
```

```{r}
correlations <- cor(segData)
dim(correlations)
```
```{r}
correlations[1:5, 1:5]
```

## Correlations between variables
```{r}
corrplot(correlations,
         order = "hclust")
```

FILTER BASED ON CORRELATIONS:

```{r}
highCorr <- findCorrelation(correlations, cutoff = .75)
length(highCorr)
```
```{r}
head(highCorr)
```

```{r}
filteredSegData <- segData[, -highCorr]
```
