---
  title: "Workshop 02"
author: "Natalia Khapaeva"
date: "02 12 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## What are we doing here?
- Using visualizations, explore the predictor variables to understand their distributions as well as the relationships between predictors.
- Do there appear to be any outliers in the data? Are any predictors skewed?
- Are there any relevant transformations of one or more predictors that might improve the classification model?

## Package Installation
### Caret
The Caret package has some issues with versions compatibility, so the installation is a bit tricky.
```{r eval = FALSE}
install.packages('caret',dependencies=TRUE)
#DO NOT RUN NOW
update.packages(checkBuilt = TRUE)
```

### Other packages
```{r eval = FALSE}
install.packages("corrplot")
install.packages("e1071")
install.packages("lattice")
install.packages("Amelia")
install.packages("epade")
install.packages("mlbench")
```

```{r}
library(mlbench)
library(AppliedPredictiveModeling)
library(corrplot)
library(caret)
library(e1071)   
```

## About the dataset
The UC Irvine Machine Learning Repository contains a data set related to glass identification. 
The data consist of 214 glass samples labeled as one of seven class categories. 
There are nine predictors, including the refractive index and percentages of eight elements: Na, Mg, Al, Si, K, Ca, Ba, and Fe.

```{r}
data(Glass)
structure(Glass)
```
```{r}
summary(Glass)
```

### Data preparation
Remove 10th collumn
```{r}
glass_cut <- Glass[, -(10)]
```
Split to predictors and classes
```{r}
classes <- Glass$Type
predictors <- glass_cut[, !names(Glass) %in% c("Type")]
```

## Tasks
### Using visualizations, explore the predictor variables to understand their distributions as well as the relationships between predictors.

Distribution of all predictors
```{r}
par(mfrow=c(3,3))
for (name in names(predictors)) {
  hist(predictors[[name]], breaks = 50, main = paste("Histogram of", name), xlab = name)
}
```
Correlation of predictors
```{r}
correlations <- cor(predictors)
corrplot(correlations)
```

The highest correlated predictor is Ca
```{r}
highCorr <- findCorrelation(correlations, cutoff = .8)
paste("Highly correlated predictors:", names(predictors)[highCorr])
```

### Do there appear to be any outliers in the data? Are any predictors skewed?

Mg,K, Ca are highly skewed
```{r}
skew_values <- sapply(predictors, skewness)
barplot(skew_values, main = "Skewness of predictors", xlab = "Predictors", ylab = "Skewness")
```
### Are there any relevant transformations of one or more predictors that might improve the classification model?

BoxCox to resolve skewness of predictors
```{r}
transformation <- preProcess(predictors, method = c("BoxCox", "center", "scale", "spatialSign"))
print(transformation)
```
```{r}
print(paste("Box-Cox applied to:", paste(transformation$method$BoxCox, collapse = ", ")))
```
```{r}
transformed <- predict(transformation, glass_cut)
```

```{r}
skew_values_transformed <- sapply(transformed, skewness)
skew_together <- rbind(skew_values, skew_values_transformed)
rownames(skew_together) <- c("original", "transformed")
barplot(skew_together,
        main = "Skewness of original predictors vs transformed predictors",
        xlab = "Predictors",
        ylab = "Skewness",
        col=c("green","blue"),
        legend = rownames(skew_together), beside=TRUE)
```
